---
title: 'DS 2020 - Summer 2026 - Homework #1'
author: "Fioni Thakker"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instruction Lab Assignment

If this is the first time you use git on your machine, you should start with [chapter 7](https://happygitwithr.com/hello-git)

Follow the step-by-step instructions in [Chapter 12 of Happy git with R](https://happygitwithr.com/rstudio-git-github.html), create your own **public** repo under your account. Note: don't do the last step. DO NOT delete the repo!

## part 1

Describe your experience: make a note of each step. In case you run into a problem, describe it and the error message.

> please write your experience below

My experience:

I already have a fair share of experience with Git, so this assignment was mostly a basic refresher for me. However, it was still good to practice the workflow and get the hang of things again!

Step 1: I opened my terminal and initialized Git tracking locally in my project folder.

Step 2: I linked my local directory to my existing repository hosted on GitHub.

Step 3: I successfully practiced the workflow cycle by tracking files, committing the updates locally, syncing changes, and pushing them to the cloud. Everything works smoothly!

## part 2

Create a new file or add any types of file into your local repository and practice the "add-commit-pull-push" cycle. Then check your remote on GitHub to make sure it succeeds.

## part 3

paste the URL of your public repo here

> paste it here: <https://github.com/fionithakker28/ds2020-fioni>

# ggplot2 Homework Assignment

## Swiss Banknotes

1.  Download the RMarkdown file with these homework instructions to use as a template for your work. Make sure to replace "Your Name" in the YAML with your name.

2.  The R package `mclust` contains a data set called `banknote`, consisting of (physical) measurements on 200 Swiss bank notes, 100 of which are genuine, while the other half is counterfeit. Load this data set (you might have to install the package) using the code below.

```{r, message = FALSE}
# install.packages("mclust")
library(mclust) # if this throws an error of the form 'there is no package called mcmclust', uncomment the line above, run it once, then comment the line out again and run the code chunk again.
library(ggplot2)
data(banknote)
```

3.  Use one of our object inspecting functions and interpret the result in the data that you see.

```{r}
str(banknote)
```

4.  Use the package `ggplot2` to draw a barchart of `Status`. Map `Status` to the fill color of the barchart.

```{r}
ggplot(data = banknote, aes(x = Status, fill = Status)) +
  geom_bar() +
  labs(title = "Count of Real vs Fake Banknotes", 
       y = "Number of Banknotes")
```

5.  Use the package `ggplot2` to draw a histogram of one of the variables in the dataset that shows a distinction between genuine and counterfeit banknotes. Use fill color to show this difference. Choose the binwidth such that there are no gaps in the middle range of the histogram.

```{r}
ggplot(data = banknote, aes(x = Bottom, fill = Status)) +
  geom_histogram(binwidth = 0.3, position = "identity", alpha = 0.6) +
  labs(title = "Histogram of Bottom Margin Width", 
       x = "Bottom Margin", 
       y = "Count")
```

6.  Use the package `ggplot2` to draw a scatterplot of two (continuous) measurements, color by `Status`. Find a pair of measurements that allow you to separate perfectly between genuine and counterfeit banknotes.

```{r}
ggplot(data = banknote, aes(x = Diagonal, y = Top, color = Status)) +
  geom_point() +
  labs(title = "Banknote Separation by Diagonal and Top Margin", 
       x = "Diagonal Length", 
       y = "Top Margin Width")
```

7.  For each of the three figures above, write a three sentence summary, describing the

    a.  **structure of the plot**: what type of plot is it? Which variables are mapped to x, to y, and to the (fill) color?

- **Figure 1 (Bar Chart):** This is a simple bar chart. The x-axis shows if the banknote is counterfeit or genuine, the y-axis shows the count of notes, and the bars are filled with different colors based on the status.
- **Figure 2 (Histogram):** This is a histogram. The x-axis shows the bottom margin measurements, the y-axis shows the count, and the bars use two different colors to separate the real and fake notes.
- **Figure 3 (Scatterplot):** This is a scatterplot. The x-axis shows the diagonal length, the y-axis shows the top margin width, and the dots are colored red for counterfeit and blue for genuine notes.

```{r}
  
```

b.  **main message of the plot**: what is your main finding, i.e. what do you want viewers to learn from the plot?

- **Figure 1 (Bar Chart):** The main message is that the dataset is perfectly balanced. It shows we have exactly 100 counterfeit notes and 100 genuine notes.
- **Figure 2 (Histogram):** The main finding is that counterfeit notes generally have a much larger bottom margin measurement than genuine notes.
- **Figure 3 (Scatterplot):** The main takeaway is that these two measurements separate the notes perfectly. The fake notes stay in one clean group on the left, and the real notes form a completely separate group on the right.

```{r}

```

c.  **additional message**: point out anomalies or outliers, if there are any.

- **Figure 1 (Bar Chart):** There are no anomalies here because both bars are exactly the same height.
- **Figure 2 (Histogram):** There is a small area in the middle where the real and fake notes mix together a little bit, but they still form two separate hills.
- **Figure 3 (Scatterplot):** There is one major anomaly where a single blue dot (a real note) is sitting completely inside the red group of fake notes. There is also another blue dot that sits way lower than the rest of its group.

```{r}

```

<!-- -->

8.  Ensure that your submission is reproducible. We will load your Rmd file into RStudio and knit it. This should match your submitted html file

For the submission: submit your solution in an R Markdown file. Also submit the corresponding html file with it.
